The Gillespie Algorithm: Stochastic Modeling of Molecular Reaction Networks

### 2024-11-27 paper.Rmd
update applicat8ion

# Application {#sec-application}
## Photodimerization Simulation
We can simulate the dimerization of DNA pyrimidines, a molecular reaction which can lead to skin cancer. Bipyrimidine photodimerization occurs when UV radiation alters the chemical bonds of two consecutive pyrimidine bases, with two possible outcomes: the formation of the cyclobutene pyrimidine dimer (CPD) if a cycloaddition between the C5-C6 double bond of the pyrimidines occurs, or the formation of pyrimidine (6-4) pyrimidone (64-PP) if  a covalent bond forms between C4 and C6 of the pyrimidines (@MartinezFernandez2022).

@@

## Radioactive Decay Simulation

>>>>

# Application {#sec-application}

This section explores the practical application of the Gillespie Algorithm by simulating two molecular systems. The first simulation, [Photodimerization Simulation](#sec-Photodimerization), models the photodimerization of thymine, a molecular reaction relevant to DNA damage caused by ultraviolet (UV) radiation. The second simulation, [Radioactive Decay Simulation](#sec-Decay), examines the stochastic decay of radioactive isotopes, a fundamental process in nuclear physics. Both simulations references the `GillespieSSA` package (@GillespieSSA), a R implementation of the Gillespie Algorithm that facilitates the generation of statistically accurate trajectories for molecular reaction systems. By consulting this package, we can simulate complex stochastic processes and explore the dynamics of these reaction systems.

## Photodimerization Simulation {#sec-Photodimerization}
We can simulate the dimerization of DNA pyrimidines, a molecular reaction which can lead to skin cancer. Bipyrimidine photodimerization occurs when UV radiation alters the chemical bonds of two consecutive pyrimidine bases, with two possible outcomes: the formation of the cyclobutene pyrimidine dimer (CPD) if a cycloaddition between the C5-C6 double bond of the pyrimidines occurs, or the formation of pyrimidine (6-4) pyrimidone (64-PP) if  a covalent bond forms between C4 and C6 of the pyrimidines (@MartinezFernandez2022).

@@

## Radioactive Decay Simulation {#sec-Decay}
